Ground state energy of quantum dots using the coupled cluster method

### 2018-06-13 project_2/article/article.tex
Start on review of results

% It's nice to be able to write your own name
\usepackage[T1]{fontenc}
% Automatic clickable links
\usepackage{hyperref}
% SI-units
\usepackage[binary-units]{siunitx}
% Enhanced math formatting
\usepackage{amsmath}
% Extended math symbols
\usepackage{amssymb}
% Include proof environment
\usepackage{amsthm}
\usepackage{enumerate}
% Import the tensor package for tensors
\usepackage{tensor}
% Include dod and dpd fracs
\usepackage{commath}
% Include font for the identity operator
\usepackage{dsfont}
% Include Tikz
\usepackage{tikz}
% Tikz add-on
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
% Several figures in the same figure
\usepackage{subfig}
% Appendix
\usepackage[toc, page]{appendix}
\usepackage{simpler-wick}
% Code listings
\usepackage{listings, lstautogobble}
\usepackage{minted}
% Colors
\usepackage{xcolor}
% Directory tree package
\usepackage{dirtree, varwidth}
% Multirow (for table)
\usepackage{multirow}

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{Makefile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.  To
        ensure that all requirements of the program are satisfied you can
        install with \mintinline{bash}{make installr} instead. Running
        \mintinline{bash}{py.test} on the command-line in the
        top-directory\footnote{That is, in
        \texttt{FYS4411/project\_2/coupled-cluster}.} will run the test scripts
        located in the \texttt{tests}-directory to make sure that the
        installation passed successfully.

@@

        The first subsection, \mintinline{bash}{matrix_elemtents},
        contains methods for computing matrix elements in harmonic
        oscillator basis from an analytical
        expression\cite{anisimovas1998energy}.  Computing the matrix
        elements is a very intensive task, and the central functions are
        therefore implemented in C++ with a Cython interface to enable use in
        Python\footnote{The C++ code is inherited from Alocias Mariadason from
        his project in FYS4411 last year with slight modifications. The original
        source code can be found here
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.cpp}
        and
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.hpp}.
        }.

        The \mintinline{bash}{hartree_fock} subsection contains methods
        for changing from harmonic oscillator basis to Hartree-Fock basis as
        well as an implementation of the self-consistent-field algorithm to
        make this transaction possible.

        The \mintinline{bash}{schemes} subsection is arguably the most important
        part of this project. The subsection has three different classes that
        perform the exact same computations, but in different and increasingly
        intelligent ways. First, \mintinline{python}{CoupledClusterDoubles} is
        the most straightforward and naïve way to solve the CCD amplitude
        equations.  Second, because an overbearing amount of the elements in the
        operator matrices in this problem are zero, we have implemented a sparse
        matrix CCD solver in \mintinline{python}{CoupledClusterDoublesSparse}.
        Third, \mintinline{python}{CoupledClusterDoublesOptimized} is
        parallellized and optimized with memory use and number of floating point
        operations in mind.

@@

    \subsection{Implementation details}
        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
        \mintinline{python}{CoupledClusterDoublesOptimized} we make use of the
        intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}.  In the optimized class special care was taken
        to ensure that every contraction was computed in an optimal way, with
        the minimization of memory use and CPU time as the overall goal.
        Moreover, several methods in NumPy for doing linear algebra was tested,
        including but not limited to \mintinline{python}{numpy.einsum},
        \mintinline{python}{numpy.tensordot}, and reshaping datastructures to
        two dimensions and using \mintinline{python}{numpy.matmul}.
        Additionally, we tried to make use of the python package
        \mintinline{python}{dask} which allows for arrays that are larger than
        the computer's memory buffer as well as automated concurrency. We found
        that the dask arrays was a bad fit for our code base, but we ended up
        using \mintinline{python}{numba} for paralellization of some
        iterative loops as well as just-in-time compilation of some functions.

@@

        The simplest way to try to ``massage'' convergence out of the CCD-method
        is to use \emph{damping} where you include a part of the result from the
        previous iteration i.e.,
        \begin{align}
            \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
            \label{eq:mixing}
        \end{align}
        where $t^{k + 1}$ is the current value computed using
        \autoref{eq:iterative_amplitude} and $t^k$ is the previous value for the
        amplitude. Choosing $\theta \in [0, 1]$ we can tune how much of the
        previous amplitude we wish to include in the new state. This allows for
        a more gradual transition between the iterations. We now use
        $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.

@@

        All the results from our simulations are displayed in tables \ref{tab:N2},
        \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. They show the ground state
        energy as a result of restricted Hartree-Fock SCF iterations, and coupled
        cluster doubles with harmonic oscillator basis and Hatree-Fock basis for
        systems of increasing size, $N=2$, $N=6$, $N=12$ and $N=20$ particles
        respectively. The results are thoroughtly discussed in the next section.

@@

        \begin{table}
            \centering
            \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
            not achieve convergence using the harmonic oscillator basis for the
            lower frequency values and large number of shells. No convergence is
            marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
                          & $4$ & $43.663267$ & $45.837079$ & $43.309845$ \\
                          & $5$ & $41.108851$ & $45.456883$ & $40.654710$ \\
                          & $6$ & $40.750512$ & \nan & $40.068340$ \\
                    $0.5$ & $7$ & $40.302719$ & \nan & $39.508500$ \\
                          & $8$ & $40.263752$ & \nan & $39.399128$ \\
                          & $9$ & $40.216688$ & \nan & $39.329311$ \\
                          & $10$ & $40.216252$ & \nan & $39.309409$ \\
                          & $11$ & $40.216195$ & \nan & $39.296007$ \\
                          & $12$ & $40.216165$ & \nan & $39.285968$ \\
                    \hline
                          & $3$ & $73.765549$ & $73.765549$ & $73.765549$ \\
                          & $4$ & $70.673849$ & $73.314476$ & $70.324250$ \\
                          & $5$ & $67.569930$ & $72.990679$ & $67.031096$ \\
                          & $6$ & $67.296869$ & \nan & $66.526677$ \\
                    $1.0$ & $7$ & $66.934745$ & \nan & $66.049564$ \\
                          & $8$ & $66.923094$ & \nan & $65.972157$ \\
                          & $9$ & $66.912244$ & \nan & $65.921205$ \\
                          & $10$ & $66.912035$ & \nan & $65.889281$ \\
                          & $11$ & $66.911365$ & \nan & $65.866715$ \\
                          & $12$ & $66.911364$ & \nan & $65.849776$ \\
                    \hline
                          & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
                          & $4$ & $117.339642$ & $120.296556$ & $116.995036$ \\
                          & $5$ & $113.660396$ & $120.007146$ & $113.048934$ \\
                          & $6$ & $113.484866$ & $119.759037$ & $112.658821$ \\
                    $2.0$ & $7$ & $113.247601$ & $119.662199$ & $112.309482$ \\
                          & $8$ & $113.246579$ & $119.584733$ & $112.235521$ \\
                          & $9$ & $113.246303$ & $119.524394$ & $112.181828$ \\
                          & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
                          & $11$ & $113.245256$ & $119.430353$ & $112.109973$ \\
                          & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
                    \hline
                          & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
                          & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
                          & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
                          & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
                    $5.0$ & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
                          & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
                          & $9$ & $234.190797$ & $241.221056$ & $233.070205$ \\
                          & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
                          & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
                          & $12$ & $234.190553$ & $241.068091$ & $232.948528$
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N12}
        \end{table}

        \begin{table}
            \centering
            \caption{In this table we look at $N = 20$ particles ($R = 4$
            shells).  We did not achieve convergence using the harmonic
            oscillator basis for the lower frequency values and large number of
            shells. No convergence is marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
                          & $5$ & $168.792442$ & $177.206536$ & $168.459124$ \\
                          & $6$ & $161.339721$ & \nan & $160.594507$ \\
                          & $7$ & $159.958722$ & \nan & $158.841120$ \\
                    $1.0$ & $8$ & $158.400172$ & \nan & $157.038330$ \\
                          & $9$ & $158.226030$ & \nan & $156.676039$ \\
                          & $10$ & $158.017667$ & \nan & $156.367930$ \\
                          & $11$ & $158.010276$ & \nan & $156.292422$ \\
                          & $12$ & $158.004951$ & \nan & $156.238258$ \\
                    \hline
                          & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
                          & $5$ & $276.898196$ & $286.159148$ & $276.381708$ \\
                          & $6$ & $267.269712$ & $285.614958$ & $266.413122$ \\
                          & $7$ & $266.213200$ & \nan & $264.969415$ \\
                    $2.0$ & $8$ & $264.933622$ & \nan & $263.434546$ \\
                          & $9$ & $264.874009$ & \nan & $263.215451$ \\
                          & $10$ & $264.809954$ & \nan & $263.046195$ \\
                          & $11$ & $264.809901$ & \nan & $262.963703$ \\
                          & $12$ & $264.809306$ & \nan & $262.899698$ \\
                    \hline
                          & $4$ & $563.773952$ & $563.773952$ & $563.773952$ \\
                          & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
                          & $6$ & $540.804720$ & $562.692231$ & $539.824400$ \\
                          & $7$ & $540.227793$ & $562.306123$ & $538.886074$ \\
                    $5.0$ & $8$ & $539.499326$ & $562.114279$ & $537.925127$ \\
                          & $9$ & $539.495941$ & \nan & $537.769045$ \\
                          & $10$ & $539.494611$ & \nan & $537.646668$ \\
                          & $11$ & $539.493513$ & \nan & $537.548828$ \\
                          & $12$ & $539.491764$ & \nan & $537.470616$ \\
                    \hline
                           & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
                           & $5$ & $961.371081$ & $972.439478$ & $960.862053$ \\
                           & $6$ & $948.057077$ & $972.002302$ & $947.019789$ \\
                           & $7$ & $947.765474$ & $971.716015$ & $946.399546$ \\
                    $10.0$ & $8$ & $947.410305$ & $971.508584$ & $945.827806$ \\
                           & $9$ & $947.409440$ & $971.332133$ & $945.663820$ \\
                           & $10$ & $947.404930$ & $971.193592$ & $945.528489$ \\
                           & $11$ & $947.404361$ & $971.076617$ & $945.424175$ \\
                           & $12$ & $947.403875$ & $970.978571$ & $945.339080$ \\
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N20}
        \end{table}

    \subsection{Running time}
        Here we show how the three implementations of the CCD-method scales as a
        function of the shell number $R$ for $N = 2$ and $N = 6$, in table
        \ref{fig:running_time_2} and \ref{fig:running_time_6} respectively. The same
        behaviour is shown for higher number of particles. In \autoref{fig:running_time_6_zoom}
        we have plotted the computation time for $N=6$ particles again, but we
        have excluded the naïve method for better comparison of the sparse and
        optimized method.

@@

    \subsection{Performance comparison}
        Somewhat surprisingly, the Hartree-Fock method outperforms the coupled
        cluster doubles when the harmonic oscillator basis is used. This
        happens for almost any configuration except the smallest system
        with $N=2$ particles. This comes from the fact that the HO basis is
        not very good at representing the states in the system. As we have
        already stated, because the Hartree-Fock SCF iterations is a variational
        methdod, the basis is varied in order to find the minimal energy. The new
        basis states will be linear combinations of the old basis states. When
        the iterations are finished, the new transformed HF basis will then be
        much better at representing the possible quantum configurations compared
        to the HO basis. This effect is only apparent in system with a larger
        number of particles and/or low frequency. This is because in a system with
        a large number of particles relative to the frequency particle-particle
        interaction will be more prominent than the harmonic oscillator potential
        the particles are subject to. I other words, the HO basis would fit rather
        well.

@@

        The rest of results have been compared with the master thesis of M. P.
        Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
        energies, one set from RHF and one set from CCSD code with harmonic
        oscillator basis functions. Our CCD code with Hartree-Fock basis will
        for some configurations\footnote{By configurations we mean frequency
        $\omega$ and number of particles $N$.} beat CCSD with plain harmonic
        oscillator basis. For 12 shells we can compare with the results from M.
        P. Lohne et al.\cite{lohne2011ab}, but in this article an
        \emph{effective interaction} for the Hamiltonian with a CCSD code has
        been used. These results are therefore significanlty better than ours,
        but we can get a ``ball-park'' idea to benchmark against.

>>>>

% It's nice to be able to write your own name
\usepackage[T1]{fontenc}
% Automatic clickable links
\usepackage{hyperref}
% SI-units
\usepackage[binary-units]{siunitx}
% Enhanced math formatting
\usepackage{amsmath}
% Extended math symbols
\usepackage{amssymb}
% Include proof environment
\usepackage{amsthm}
\usepackage{enumerate}
% Import the tensor package for tensors
\usepackage{tensor}
% Include dod and dpd fracs
\usepackage{commath}
% Include font for the identity operator
\usepackage{dsfont}
% Include Tikz
\usepackage{tikz}
% Tikz add-on
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
% Several figures in the same figure
\usepackage{subfig}
% Appendix
\usepackage[toc, page]{appendix}
\usepackage{simpler-wick}
% Code listings
\usepackage{listings, lstautogobble}
%\usepackage{minted}
% Colors
\usepackage{xcolor}
% Directory tree package
\usepackage{dirtree, varwidth}
% Multirow (for table)
\usepackage{multirow}

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{Makefile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.  To
        ensure that all requirements of the program are satisfied you can
        install with \texttt{make installr} instead. Running
        \texttt{py.test} on the command-line in the
        top-directory\footnote{That is, in
        \texttt{FYS4411/project\_2/coupled-cluster}.} will run the test scripts
        located in the \texttt{tests}-directory to make sure that the
        installation passed successfully.

@@

        The first subsection, \texttt{matrix\_elements},
        contains methods for computing matrix elements in harmonic
        oscillator basis from an analytical
        expression\cite{anisimovas1998energy}.  Computing the matrix
        elements is a very intensive task, and the central functions are
        therefore implemented in C++ with a Cython interface to enable use in
        Python\footnote{The C++ code is inherited from Alocias Mariadason from
        his project in FYS4411 last year with slight modifications. The original
        source code can be found here
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.cpp}
        and
        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.hpp}.
        }.

        The \texttt{hartree\_fock} subsection contains methods
        for changing from harmonic oscillator basis to Hartree-Fock basis as
        well as an implementation of the self-consistent-field algorithm to
        make this transaction possible.

        The \texttt{schemes} subsection is arguably the most important
        part of this project. The subsection has three different classes that
        perform the exact same computations, but in different and increasingly
        intelligent ways. First, \texttt{CoupledClusterDoubles} is
        the most straightforward and naïve way to solve the CCD amplitude
        equations.  Second, because an overbearing amount of the elements in the
        operator matrices in this problem are zero, we have implemented a sparse
        matrix CCD solver in \texttt{CoupledClusterDoublesSparse}.
        Third, \texttt{CoupledClusterDoublesOptimized} is
        parallellized and optimized with memory use and number of floating point
        operations in mind.

@@

    \subsection{Implementation details}
        Both in \texttt{CoupledClusterDoublesSparse} and
        \texttt{CoupledClusterDoublesOptimized} we make use of the
        intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}.  In the optimized class special care was taken
        to ensure that every contraction was computed in an optimal way, with
        the minimization of memory use and CPU time as the overall goal.
        Moreover, several methods in NumPy for doing linear algebra was tested,
        including but not limited to \texttt{numpy.einsum},
        \texttt{numpy.tensordot}, and reshaping datastructures to
        two dimensions and using \texttt{numpy.matmul}.
        Additionally, we tried to make use of the python package
        \texttt{dask} which allows for arrays that are larger than
        the computer's memory buffer as well as automated concurrency. We found
        that the dask arrays was a bad fit for our code base, but we ended up
        using \texttt{numba} for paralellization of some
        iterative loops as well as just-in-time compilation of some functions.

@@

        The simplest way to try to ``massage'' convergence out of the CCD-method
        is to use \emph{damping} where you include a part of the result from the
        previous iteration i.e.,
        \begin{align}
            \tilde{t}^{k + 1} = (1 - \theta) t^{k + 1} + \theta t^k,
            \label{eq:mixing}
        \end{align}
        where $t^{k + 1}$ is the current value computed using
        \autoref{eq:iterative_amplitude} and $t^k$ is the previous value for the
        amplitude. Choosing $\theta \in [0, 1]$ we can tune how much of the
        previous amplitude we wish to include in the new state. This allows for
        a more gradual transition between the iterations. We now use
        $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.

@@

        All the results from our simulations are displayed in tables
        \ref{tab:N2}, \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. The tables
        show the results for systems of increasing size, $N=2$, $N=6$, $N=12$
        and $N=20$ particles respectively. The results are thoroughtly discussed
        in the next section.

@@

        \begin{table}
            \centering
            \caption{Here we look at $N = 12$ particles.  We did not achieve
            convergence using the harmonic oscillator basis for the lower
            frequency values and large number of shells. No convergence is
            marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
                          & $4$ & $43.663267$ & $45.837079$ & $43.309845$ \\
                          & $5$ & $41.108851$ & $45.456883$ & $40.654710$ \\
                          & $6$ & $40.750512$ & \nan & $40.068340$ \\
                    $0.5$ & $7$ & $40.302719$ & \nan & $39.508500$ \\
                          & $8$ & $40.263752$ & \nan & $39.399128$ \\
                          & $9$ & $40.216688$ & \nan & $39.329311$ \\
                          & $10$ & $40.216252$ & \nan & $39.309409$ \\
                          & $11$ & $40.216195$ & \nan & $39.296007$ \\
                          & $12$ & $40.216165$ & \nan & $39.285968$ \\
                    \hline
                          & $3$ & $73.765549$ & $73.765549$ & $73.765549$ \\
                          & $4$ & $70.673849$ & $73.314476$ & $70.324250$ \\
                          & $5$ & $67.569930$ & $72.990679$ & $67.031096$ \\
                          & $6$ & $67.296869$ & \nan & $66.526677$ \\
                    $1.0$ & $7$ & $66.934745$ & \nan & $66.049564$ \\
                          & $8$ & $66.923094$ & \nan & $65.972157$ \\
                          & $9$ & $66.912244$ & \nan & $65.921205$ \\
                          & $10$ & $66.912035$ & \nan & $65.889281$ \\
                          & $11$ & $66.911365$ & \nan & $65.866715$ \\
                          & $12$ & $66.911364$ & \nan & $65.849776$ \\
                    \hline
                          & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
                          & $4$ & $117.339642$ & $120.296556$ & $116.995036$ \\
                          & $5$ & $113.660396$ & $120.007146$ & $113.048934$ \\
                          & $6$ & $113.484866$ & $119.759037$ & $112.658821$ \\
                    $2.0$ & $7$ & $113.247601$ & $119.662199$ & $112.309482$ \\
                          & $8$ & $113.246579$ & $119.584733$ & $112.235521$ \\
                          & $9$ & $113.246303$ & $119.524394$ & $112.181828$ \\
                          & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
                          & $11$ & $113.245256$ & $119.430353$ & $112.109973$ \\
                          & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
                    \hline
                          & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
                          & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
                          & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
                          & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
                    $5.0$ & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
                          & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
                          & $9$ & $234.190797$ & $241.221056$ & $233.070205$ \\
                          & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
                          & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
                          & $12$ & $234.190553$ & $241.068091$ & $232.948528$
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N12}
        \end{table}

        \begin{table}
            \centering
            \caption{In this table we look at $N = 20$ particles.  We did not
            achieve convergence using the harmonic oscillator basis for the
            lower frequency values and large number of shells. No convergence is
            marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
                          & $5$ & $168.792442$ & $177.206536$ & $168.459124$ \\
                          & $6$ & $161.339721$ & \nan & $160.594507$ \\
                          & $7$ & $159.958722$ & \nan & $158.841120$ \\
                    $1.0$ & $8$ & $158.400172$ & \nan & $157.038330$ \\
                          & $9$ & $158.226030$ & \nan & $156.676039$ \\
                          & $10$ & $158.017667$ & \nan & $156.367930$ \\
                          & $11$ & $158.010276$ & \nan & $156.292422$ \\
                          & $12$ & $158.004951$ & \nan & $156.238258$ \\
                    \hline
                          & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
                          & $5$ & $276.898196$ & $286.159148$ & $276.381708$ \\
                          & $6$ & $267.269712$ & $285.614958$ & $266.413122$ \\
                          & $7$ & $266.213200$ & \nan & $264.969415$ \\
                    $2.0$ & $8$ & $264.933622$ & \nan & $263.434546$ \\
                          & $9$ & $264.874009$ & \nan & $263.215451$ \\
                          & $10$ & $264.809954$ & \nan & $263.046195$ \\
                          & $11$ & $264.809901$ & \nan & $262.963703$ \\
                          & $12$ & $264.809306$ & \nan & $262.899698$ \\
                    \hline
                          & $4$ & $563.773952$ & $563.773952$ & $563.773952$ \\
                          & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
                          & $6$ & $540.804720$ & $562.692231$ & $539.824400$ \\
                          & $7$ & $540.227793$ & $562.306123$ & $538.886074$ \\
                    $5.0$ & $8$ & $539.499326$ & $562.114279$ & $537.925127$ \\
                          & $9$ & $539.495941$ & \nan & $537.769045$ \\
                          & $10$ & $539.494611$ & \nan & $537.646668$ \\
                          & $11$ & $539.493513$ & \nan & $537.548828$ \\
                          & $12$ & $539.491764$ & \nan & $537.470616$ \\
                    \hline
                           & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
                           & $5$ & $961.371081$ & $972.439478$ & $960.862053$ \\
                           & $6$ & $948.057077$ & $972.002302$ & $947.019789$ \\
                           & $7$ & $947.765474$ & $971.716015$ & $946.399546$ \\
                    $10.0$ & $8$ & $947.410305$ & $971.508584$ & $945.827806$ \\
                           & $9$ & $947.409440$ & $971.332133$ & $945.663820$ \\
                           & $10$ & $947.404930$ & $971.193592$ & $945.528489$ \\
                           & $11$ & $947.404361$ & $971.076617$ & $945.424175$ \\
                           & $12$ & $947.403875$ & $970.978571$ & $945.339080$ \\
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N20}
        \end{table}

    \subsection{Running time}
        Here we show how the three implementations of the CCD-method scales as a
        function of the shell number $R$ for $N = 2$ and $N = 6$, in figures
        \ref{fig:running_time_2} and \ref{fig:running_time_6} respectively. The
        same behaviour is shown for higher number of particles. In
        \autoref{fig:running_time_6_zoom} we have plotted the computation time
        for $N=6$ particles again, but we have excluded the naïve method for
        better comparison of the sparse and optimized method.

@@

    \subsection{Performance comparison}
        Somewhat surprisingly, the Hartree-Fock method outperforms the coupled
        cluster doubles when the harmonic oscillator basis is used. This happens
        for almost any configuration except the smallest system with $N=2$
        particles. This comes from the fact that the HO basis is not very good
        at representing the states in the system. Because the Hartree-Fock SCF
        iterations is a variational method, the basis is varied in order to find
        the minimal energy. The new basis states will be linear combinations of
        the old basis states. When the iterations are finished, the new
        transformed HF basis will then be much better at representing the
        possible quantum configurations compared to the HO basis. This effect is
        only apparent in systems with a larger number of particles and/or low
        frequency. This is because in a system with a large number of particles
        relative to the frequency will experience that the particle-particle
        interaction will be more prominent than the harmonic oscillator
        potential the particles are subject to. The trapped
        particles in the harmonic oscillator potential will be pushed out of the
        well unless we increase the frequency thus increasing the strength of
        the trap.

        In other words, our first guess using the harmonic oscillator basis
        functions as an approximation to the exact eigenfunction of the quantum
        dot Hamiltonian is only a good first attempt when the number of
        particles is low and/or the frequency is very high. In those situations
        we can treat the particle-particle interaction as a small perturbation.

@@

        The rest of the results have been compared with the master thesis of M.
        P.  Lohne\cite{lohne2010coupled} for up to $R = 10$ shells. Lohne gets
        two sets of energies, one set from RHF and one set from CCSD code with
        harmonic oscillator basis functions. Our CCD code with Hartree-Fock
        basis will for some configurations\footnote{By configurations we mean
        frequency $\omega$ and number of particles $N$.} beat CCSD with plain
        harmonic oscillator basis. For $R = 12$ shells we can compare with the
        results from M. P. Lohne et al.\cite{lohne2011ab}, but in this article
        an \emph{effective interaction} for the Hamiltonian with a CCSD code
        using the HF basis has been used. These results are therefore
        significantly better than ours, but we can get a ``ball-park'' idea to
        benchmark against.
